---
title: 'Individual Research Project - Histogram Visualization Tutorial'
author: "Jonathan Schierbaum"
date: "FALL 2020"
output:
  html_document: default
  pdf_document: default
highlight: pygments
---
<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, tidy = FALSE, size = "small", message = FALSE, fig.align = 'center') 
options(width = 999)

packages <- c("knitr", "dplyr", "lubridate", 
              "reshape2", "ggplot2", "ggnewscale", 
              "gganimate")

missing <- packages[!(packages %in% installed.packages())]

if(length(missing) > 0) install.packages(missing)

invisible(suppressMessages(lapply(packages, library, character.only = TRUE)))

rm(packages, missing)
```

# Sharpening Your Histogram Game

Histograms are often cited as the most used visualization by the statistics community.

I am going to show you a few simple techniques that you can implement easily to maximize a histogram's potential for interpretation.


## Data description...it's a bit strange!

**Read about the data**
[link](https://www.google.com/covid19/mobility/)

**Download the data, aggregated for you**
[link](https://www.kaggle.com/douglaskgaraujo/opportunity-insights-real-time-economic-tracker-us)


**The data we're using**
Measures the change in time spent at home, by day.

Strangely, each day is compared against the median value for that day of the week the 5‑week period Jan 3 – Feb 6, 2020. 

**Because people already spend much of the day at places of residence (even on workdays), the capacity for change isn’t so large.**

For example: 

In a given day during pandemic, let's say a hypothetical person before the pandemic spent 9 hours outside the home, and 15 hours in the home. 

During the pandemic, If they then spent their entire lives at home, their time in residence would only increase by 60%.


**Complicated, difficult to interpret dataset! Histograms can still help!**

### Load data

```{r load_data}
fp_raw_data <- "~/GitHub/exercise-histogram-visualization/raw_data/"

fp_figures <- "~/GitHub/exercise-histogram-visualization/figures/"


# loading the data in
data1   <- read.csv(paste0(fp_raw_data, "Google Mobility - City - Daily.csv"), 
                    stringsAsFactors=F)                                         # stringsAsFactors=FALSE best for beginners
                                                                                # to set their own factors.
cityid  <- read.csv(paste0(fp_raw_data, "GeoIDs - City.csv"), 
                    stringsAsFactors=F)                                         

data2   <- read.csv(paste0(fp_raw_data, "Google Mobility - State - Daily.csv"), 
                    stringsAsFactors=F)  

stateid <- read.csv(paste0(fp_raw_data, "GeoIDs - State.csv"), 
                    stringsAsFactors=F)                                         # Geo-ID files contain identifiers we will need.


# add city and state identifiers
city_data  <- left_join(data1, cityid, by = "cityid")                           # left-join finds rows with matching entries
state_data <- left_join(data2, stateid, by = "statefips")                       # in the cityid column, and adds the 2nd dataframe's
                                                                                # columns to the first
 
# make the date more manageable
city_data  <- mutate(city_data, date = make_date(year, month, day))             # using lubridate to make a new, properly coded date.
city_data  <- mutate(city_data, month = month(date, label = T))                 # using that new date to make new month and day columns.
city_data  <- mutate(city_data, weekday = wday(date, label = T))                # label = T adds abbreviations for days instead of 1,2,3...

state_data <- mutate(state_data, date = make_date(year, month, day))
state_data <- mutate(state_data, month = month(date, label = T))
state_data <- mutate(state_data, weekday = wday(date, label = T))               


# remove weekends
city_wk     <- filter(city_data,                                                # filter() + between() is very useful
                      between(wday(date), 2, 6))                                # wday() returns day of week, between() keeps weekdays

city_wkend  <- filter(city_data,                                                # filter weekends to separate dataframe
                      wday(date) == 1 |wday(date) == 7)

state_wk    <- filter(state_data,                                               # repeat for state data.
                      between(wday(date), 2, 6))

state_wkend <- filter(state_data, 
                      wday(date) == 1 |wday(date) == 7)

# exploratory plot                                                                  # using ggplot2 is like using building blocks

png(file=paste0(fp_figures, "scatterplot_all_data", ".png"), width = 8, height = 5, unit = "in", res = 225, antialias = "cleartype")

ggplot(data = state_wk, aes(date, gps_residential)) +                               # the + sign adds another element
  scale_y_continuous(labels = scales::percent) +                                    # your plot not coming up? check your chain of 
  geom_point() +                                                                    # + signs.
                                                                                    # scale_y_continuous(labels = scales::percent)
  ggtitle("Daily Hours at Home Across All States, Feb - Sep '20",                   # changes our axis to percent.
        subtitle = "each data point is a statewide weekday average (by day), compared to january '20 median levels" ) +  
  xlab("Month of Record") +                                                         
  ylab("Daily Changes in Hours Home (Day of Week vs. Jan Median)")

dev.off()
```

## Histogram introduction

```{r hist_intro}
# Let's start with a look at cities.

# make city data frame, filter to weeks of interest
city_target <- filter(city_wk,                                                  # filtering weeks 11-14 
                      between(isoweek(date), 11,14))                            # wrapping date in isoweek() returns week of year.
                                                                                # I use isoweek() because its weeks start on monday

# double check for filter errors
length(unique(city_target$date))                                                # expecting 4 weeks x 5 days = 20 days
length(unique(city_target$weekday))


# week check, locate starting week
chk <- as.Date(c("2020-03-02", "2020-03-09", "2020-03-16"))                     # this data is from 3/9/2020 to 4/3/2020
isoweek(chk) 


# wow, a histogram!                                                             # to change previous plot to histogram, remove
png(file=paste0(fp_figures, "intro_histogram", ".png"), 
    width = 8, height = 5, unit = "in", res = 225, antialias = "cleartype")

ggplot(data = city_target, aes(gps_residential)) +                              # isoweek(date) from the aesthetic mapping
  geom_histogram() +                                                            # and remove geom_point() and add geom_hist()  
  ggtitle("Weekday Hours at Home Across 53 U.S. Cities in March and April 2020",
          subtitle = "each data point is a citywide weekday average (by day), compared to january '20 median levels" ) +
  
  xlab("Daily Changes in Hours Home (Each Weekday vs. Jan Median)") + 
  ylab("No. of Days"
 ) + 
  
# Add a median line  
  geom_vline(aes(xintercept=median(gps_residential)),                           # a sneak peak. geom_vline() needs its own x value mapping!
             color="red", linetype="dashed", size=1 )                           # anything we put inside aes() within a geom layer overrides
                                                                                # the original aes settings from the ggplot2 call.
dev.off()
```
### Histograms are not bar charts

Visually, they are similar. Both use bars placed side-by-side, and bar height is a main visual cue, but the small differences between them change interpretation significantly.

**Bar charts**
Shows categorical data (and sometimes time series data) on the x axis.

The visual cue for a value in a bar chart is bar height.

Primarily used to show illustrate categories.

**histograms**
Show a continuous variable on the horizontal axis. 
 
Uses area (i.e. width times height) as a visual cue.

Primarily used to show illustrate distributions.

The width of bars can vary, (i.e. to see the long-tail of a distribution)

This means you read the two chart types differently. The bar chart is for categories, and the histogram is for distributions. 

**Main note: Bar charts and histograms are different.**

```{r hist_vs_bar, fig.height=6}
# Bar charts aren't histograms! Bar chart EX; cities with highest average

avgchg <- aggregate(gps_residential ~ cityname, data=city_target, mean)         # aggregate() is grouping by city name and returning 1
avgchg <- avgchg[order(avgchg$gps_residential, decreasing=TRUE),]               # value per city. order() ranks those values. 

                                                                                # reorder() gives some structure to the plot

ggplot(data = avgchg, aes(reorder(cityname, gps_residential),                   # we're mapping the y value ourselves now (no count),
                          gps_residential)) +                                   # We must add y values inside aes()  
  geom_bar(stat="identity") +                                                   # stat = ".. makes the bar heights reflect their value 
                                                                                # within the data (used within geom)
  coord_flip() +                                                                # coord_flip() rotates the graph, would look even more like
                                                                                # a hist without this.
  ggtitle("Weekday Hours at Home Across 53 U.S. Cities in March and April 2020",               
        subtitle = "the mean value of citywide daily records between 3/9/20 to 4/3/20, compared to january '20 median levels" ) +          
  ylab("Daily Change in Hours Home (Weekday vs. Jan Median)") +  
  xlab("Select U.S. Cities")

# Moving forward
png(file=paste0(fp_figures, "bar_chart_ex", ".png"), 
    width = 9, height = 6.5, unit = "in", res = 225, antialias = "cleartype")

ggplot(data = avgchg, aes(reorder(cityname, gps_residential) , gps_residential)) +
  geom_bar(stat="identity") + 
  
  geom_hline(aes(yintercept = mean(gps_residential)),                           # Notice usage of geom_hline(). The line is 
             color="red",                                                       # horizontal until coord-flip. Also, notice the y intercept
             linetype="dashed",                                                 # function is now mean(), and not median()
             size=1) + 
  
  coord_flip() +
  
  geom_text(aes(x = "Phoenix", y = (avgchg[43,2] + 0.015)),                     # These text are just added to draw attention.
            label="Phoenix", hjust = 0.75) +                                         # Align by setting their x value within the aes() to their name
                                                                                # and their y value to their average, + a little offset.
  geom_text(aes(x = "Phoenix", y = (avgchg[43,2] + 0.015), 
            label = round(avgchg[43,2], digit = 3)), hjust = -0.75) +
  
  geom_text(aes(x = "San Diego", y = (avgchg[10,2] + 0.02)), 
            label="San Diego", hjust = 0.75) + 
  
  geom_text(aes(x = "San Diego", y = (avgchg[10,2] + 0.02), 
            label = round(avgchg[10,2], digit = 3)), hjust = -0.75) +

  ggtitle("Weekday Hours at Home Across 53 U.S. Cities in March and April 2020",               
        subtitle = "the mean value of citywide daily records between 3/9/20 to 4/3/20, compared to january '20 median levels" ) +          
  ylab("Daily Change in Hours Home (Weekday vs. Jan Median)") +  
  xlab("Select U.S. Cities")

dev.off()
```
### Transparency helps with interpretation

```{r main_hist_start}

# let's look at San Diego & Phoenix, the first 100 weekdays of the period.
yday(city_target[1,17])                                                         # yday() returns the day of year. Here, I'm looking
                                                                                # at the first record of the previous dataframe
sdph <- filter(city_wk,
               between(yday(date), 69, 208) & cityname == "San Diego" |         # filter() keeps rows that match either condition, 
               between(yday(date), 69, 208) & cityname == "Phoenix" )           # the conditions are split by | 

# basic plot
histplot <- ggplot(data = sdph, aes(gps_residential)) +                         # basic ggplot object
scale_x_continuous(labels = scales::percent) +                                  # with histograms we scale the x axis
  ggtitle("Weekday Hours at Home In Two Southwest Cities, 100 Days Of Change After 3/9/20",               
        subtitle = "100 citywide weekday averages from san diego and phoenix, compared to january '20 median levels" ) +          
        xlab("Daily Changes in Hours Home (Each Weekday vs. Jan Median)") +
        ylab("No. of Days")


# Tip #1: always add alpha
png(file=paste0(fp_figures, "histogram_ex", ".png"), 
    width = 8, height = 5, unit = "in", res = 225, antialias = "cleartype")

histplot +  geom_histogram() 

dev.off()

png(file=paste0(fp_figures, "histogram_ex_iter_1", ".png"), 
    width = 8, height = 5, unit = "in", res = 225, antialias = "cleartype")

histplot +  geom_histogram(alpha = 0.5)                                         # adding alpha at the geom level ensures that only

dev.off()                                                                       # the histogram bars will be affected.
```
### Cues that inform histogram interpretation

For a single variable...

**Shape**
Is there a skew to the left or right, or a clump in the middle?

Does it resemble a familiar distribution?

**Span**
Are there are lot of values clustered towards the maximums and minimums with nothing in between?

What happens in between the maximum value and median? Do the values cluster towards the median and quickly increase?

**Outliers**
Histograms make them very easy to spot.

## Useful histogram improvements

### Adjust bin sizes

```{r bin_sizes}

#Tip #2: adjusting bins, binwidth and breaks
histplot +  
  geom_histogram(alpha = 0.5, bins = 20) +                                      # note: bins is within geom()
  ggtitle("bins = 20")

histplot + 
  geom_histogram(alpha = 0.5, binwidth = 0.2) +                                 # note: binwidth is within geom() 
  ggtitle("binwidth = 0.2")

#mimic R plot default breaks
defaultbins <- pretty(range(sdph$gps_residential),                              # This is the default behavior of hist()
    n = nclass.Sturges(sdph$gps_residential), min.n = 1)                        # within base R, it returns a vector of breaks.

histplot + 
  geom_histogram(alpha = 0.5, breaks = defaultbins) +                           # note: breaks is within geom()
  ggtitle("pretty sturges")

#bins don't need to be equal...
weirdbins <- sample(defaultbins, 10)                                            # sample() is just being used to randomly create
                                                                                # different sized bins.
histplot + 
  geom_histogram(alpha = 0.5, breaks = weirdbins) +                             # breaks = somevector must be used for different
  ggtitle("weirdbins")                                                          # sized bins.


                                                                                
# Moving on...
histplot1 <- ggplot(data = sdph, aes(gps_residential)) +                            
scale_x_continuous(labels = scales::percent) +                                     
  ggtitle("Weekday Hours at Home In Two Southwest Cities, 100 Days Of Change After 3/9/20",               
        subtitle = "100 citywide weekday averages from san diego and phoenix, compared to january '20 median levels" ) +          
        xlab("Daily Changes in Hours Home (Each Weekday vs. Jan Median)") +
        ylab("No. of Days") +
  geom_histogram(alpha = 0.5, bins = 20)                                        # I chose 20 bins because I thought it looked best

png(file=paste0(fp_figures, "histogram_ex_iter_2", ".png"), 
    width = 8, height = 5, unit = "in", res = 225, antialias = "cleartype")

histplot1

dev.off()   
``` 


### Measures of central tendancy

```{r median_lines}

# Tip 3: add central tendency indicators (median)
histplot1 + geom_vline(aes(xintercept=median(gps_residential)),                 # geom_vline adds vertical lines.  We want this layer to
              color = "red", linetype = "dashed", size = 1 )                    # behave differently, so we make adjustments to its aes() call.
                                                                                # instead of changing its x value with every observation, we want
                                                                                # it's static position to be median(gps_residential)

# Moving on...
histplot1 <- histplot1 + 
  geom_vline(aes(xintercept=median(gps_residential)), 
              color = "red", linetype = "dashed", size = 1 )

png(file=paste0(fp_figures, "histogram_ex_iter_3", ".png"), 
    width = 8, height = 5, unit = "in", res = 225, antialias = "cleartype")

histplot1

dev.off()
``` 
### IQR ranges

```{r IQR_ranges}
 
# Tip #3 continued: central tendency indicators (IQR)
histplot1 + annotate("rect",                                          # single plot? annotate() is the easiest to use.
              xmin = quantile(sdph$gps_residential, probs = 0.25),    # Annotate takes no information from the original ggplot, so every x,y
              xmax = quantile(sdph$gps_residential, probs = 0.75),    # so every x,y must be defined.
              ymin = 0,                                               # quantile(sdph$gps_residential, probs = 0.25) returns the x value that
              ymax = 40,                                              # has 25% of the data points below it.
              alpha = .1, fill="blue")
  
histplot1 + annotate("rect", 
              xmin = quantile(sdph$gps_residential, probs = 0.10), 
              xmax = quantile(sdph$gps_residential, probs = 0.90), 
              ymin = 0,                                                         # ymin you usually want zero
              ymax = 40,                                                        # ymax is trial and error, or you can define with complex function
              alpha = .1, fill="blue")

# Can also use annotate for general highlighting   
histplot1 + annotate("rect",                                                    # for this one, I just experimented with different coordinates.
              xmin = -0.02, 
              xmax = 0.075, 
              ymin = 0, 
              ymax = 10, 
              alpha = .1, fill="yellow")
 
# Moving on...
histplot2 <- histplot1 + annotate("rect",                                                  
              xmin = quantile(sdph$gps_residential, probs = 0.25),              
              xmax = quantile(sdph$gps_residential, probs = 0.75),              
              ymin = 0,                                                         
              ymax = 40,                                                        
              alpha = .1, fill="blue")

png(file=paste0(fp_figures, "histogram_ex_iter_4", ".png"), 
    width = 8, height = 5, unit = "in", res = 225, antialias = "cleartype")

histplot2

dev.off()
```

### Rugs and color

```{r rug_and_color, fig.height=6}

# Tip 4: Rugs can be a great way to show additional distribution info 
histplot2 + geom_rug()                                                          # geom rug adds tick marks for each x value

histplot2 + geom_rug(sides = "tr",                                              # sides changes where the tick marks appear
                    outside = TRUE) +                                           # outside moves the ticks outside the gray plotting area
                    coord_cartesian(clip = "off")                               # adjusting this sometimes fixes really tiny tick marks


histplot2 + geom_rug(aes(color = cityname),                                     # moving color inside the aes() tells ggplot2 to change the
                    sides = "tr") +                                             # color of tick marks based on the value of cityname
                    coord_cartesian(clip = "off")  


# Moving on...(and adding some color)

histplot3 <- ggplot(data = sdph, aes(gps_residential,                           # Inside the aes() of the main plot, add
                                    fill = cityname,                            # fill = cityname and color = cityname.
                                    color = cityname)) +                        # ggplot will now assign colors to different  
  scale_x_continuous(                                                           # values in this column
        labels = scales::percent) +                                    
  ggtitle("Weekday Hours at Home In Two Southwest Cities, 100 Days Of Change After 3/9/20",               
        subtitle = "100 citywide weekday averages from san diego and phoenix, compared to january '20 median levels" ) +          
        xlab("Daily Changes in Hours Home (Each Weekday vs. Jan Median)") +
        ylab("No. of Days") +
  geom_histogram(
        alpha = 0.5, 
        bins = 20) +    
  geom_vline(aes(xintercept=median(gps_residential), size = "Median"),          # inside the aes of the median line, we add
        linetype = "dashed",                                                    # size = "Median" to have ggplot2 assign
        color = "red" ) +                                                       # the size legend to this geom
  scale_size_manual("", values=1,                                               # force the color and size values here.
        guide=guide_legend(override.aes = list(color=c("red"), linetype = "solid"))) +                 
  annotate("rect", fill = "blue",                                                
        xmin = quantile(sdph$gps_residential, probs = 0.25),              
        xmax = quantile(sdph$gps_residential, probs = 0.75),              
        ymin = 0,                                                         
        ymax = 40,                                                        
        alpha = .1) +
  geom_rug(aes(color = cityname),                                     
        sides = "tr") +                                             
        coord_cartesian(clip = "off") +  
        labs(fill="") +                                                         # This hides just the legend title.  Have to do it here
        labs(color="")                                                          # too, ggplot makes 2 legends, fill\color.


png(file=paste0(fp_figures, "histogram_2pop_ex", ".png"), 
    width = 8, height = 5, unit = "in", res = 225, antialias = "cleartype")

histplot3

dev.off()
```

### Grids of histograms (a.k.a. multiples, marginals, facets)

```{r dataframe_creation}
# to make it a bit more interesting, add back a few cities
cities <- filter(city_wk,                                                       # adding more 5 cities above the average, 5 cities
               between(yday(date), 69, 208) & cityname == "San Francisco" |     # below the average (based on mean bar chart above)
               between(yday(date), 69, 208) & cityname == "New York City" |
               between(yday(date), 69, 208) & cityname == "Miami" |
               between(yday(date), 69, 208) & cityname == "San Diego" |
               between(yday(date), 69, 208) & cityname == "Portland" |
               between(yday(date), 69, 208) & cityname == "Baltimore" |
               between(yday(date), 69, 208) & cityname == "Phoenix" |
               between(yday(date), 69, 208) & cityname == "Jacksonville" |
               between(yday(date), 69, 208) & cityname == "Memphis" |
               between(yday(date), 69, 208) & cityname == "Tulsa" )

# we're going to need this later...add columns for median and IQR values
cities <- group_by(cities, cityname) %>%                                        # groupby() groups the dataset by cityname
  mutate(med = median(gps_residential)) %>%                                     # mutate() then adds columns with matching entries
  mutate(IQR1 = quantile(gps_residential, probs = 0.25)) %>%                    # within groups. ie, one median for each city, repeated
  mutate(IQR2 = quantile(gps_residential, probs = 0.75))                        # in each daily record of that city.
```

```{r small_multiples, fig.width = 20, fig.height = 10}

# Tip 5: Small multiples, marginals, facets, whatever you want to call it!

ggplot(data = cities, aes(gps_residential, fill=cityname)) +                    # fill=cityname in the aes() of ggplot will pass this to
                                                                                # lower layers. 
  geom_rect(data = cities, inherit.aes = FALSE,                                 # 
     aes(xmin = IQR1, xmax = IQR2, ymin=0, ymax=40, group = cityname),          # we have to swap out annotate for geom_rect.
     fill = "#8496E4", alpha=0.01) +                                            # by mapping our new x and y columns inside its aes(), the 
                                                                                # geom is now taking its x values from our new columns. 
  geom_histogram(alpha = 0.5, bins = 20) +                                      # This is not an ideal way to highlight IQR ranges, because
                                                                                # a rectangle is drawn for each data point, causing 
                                                                                # transparency issues. That is why it is
                                                                                # moved up in the ggplot stack.

  ggtitle("Weekday Hours at Home In Ten Select Cities, 100 Days Of Change After 3/9/20",               
        subtitle = "100 citywide weekday averages from 10 select cities, compared to january '20 median levels" ) +          
        xlab("Daily Changes in Hours Home (Each Weekday vs. Jan Median)") +
        ylab("No. of Days") +
  
  
                                                                                #
  geom_vline(aes(xintercept = med, group = cityname),                           # we must map the median line's x value to the new column.
              color = "red", linetype = "dashed", size = 1 ) +                  # group = cityname tells ggplot to calculate for the group.
                                                                                # 
  facet_wrap(~ cityname, ncol=5)                                                # facet_wrap makes a new plot for each cityname. ncol is 
                                                                                # the number of columns of the graphic.


# we need this data frame of IQRs to make alpha work a bit better.
IQRs <- cities[!duplicated(cities$cityname), ]                                  # Here's a better method for making IQRs
                                                                                # we need to create a separate data frame with just 1 record
                                                                                # per city, and map it only to the geom_rect layer.
histplot4 <- ggplot(data = cities, aes(gps_residential, fill=cityname)) + 
  scale_x_continuous(labels = scales::percent) +    
  geom_histogram(alpha = 0.5, bins = 20) +
  geom_vline(aes(xintercept = med, group = cityname), 
             color = "red", linetype = "dashed", size = 1, alpha = 0.8) +
  
  geom_rect(data = IQRs, inherit.aes = FALSE,                                   # we can map a dataset to one layer only.
            aes(xmin = IQR1, xmax = IQR2, ymin=0, ymax=40, group = cityname),  
                fill = "blue", alpha=0.10) +  
  geom_rug(aes(color = cityname)) +
    ggtitle("Weekday Hours at Home In Ten Select Cities, 100 Days Of Change After 3/9/20",               
        subtitle = "100 citywide weekday averages from 10 select cities, compared to january '20 median levels" ) +          
        xlab("Daily Changes in Hours Home (Each Weekday vs. Jan Median)") +
        ylab("No. of Days") + 
  facet_wrap(~ cityname, ncol=5) 

histplot4

# I'm going to add region labels to each facet.  
# Here is the most flexible way to add text labels per facet, set up a table!
plot_text <- data.frame(
  label    = c("(Northeast)", "(South)",   "(South)", "(South)", "(Northeast)",  "(West)",  "(West)",   "(West)",    "(West)",   "(South)"),
  cityname = c("Baltimore", "Jacksonville", "Memphis", "Miami", "New York City", "Phoenix","Portland", "San Diego", "San Francisco", "Tulsa"),
  x        = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15),
  y        = c(42, 42, 42, 42, 42, 42, 42, 42, 42, 42))
  
  
  
# Let's move on, with a bit more style.  
histplot5 <- histplot4 + geom_text(data= plot_text,                             # On the left, we position by table value
                                   aes(x = x, y = y, label = label),            # The data frame we just created is mapped to the text
                                   size = 3) +                                  # layer only.
                                                                                # 
                         theme_minimal() +                                      # if you don't have time to tweak every detail,                                                                                                                         # theme_minimal() is a good bet for facets.
  
                         theme(legend.position = "none") +                      # If you want to remove a legend, position = "none"  
                 
                         theme(strip.text =                                     # Modify text of Facets.
                         element_text(face="bold", size=12,lineheight=5.0))     # You can modify everything in ggplot2. Some mods are   
                                                                                # easier than others. 

png(file=paste0(fp_figures, "histogram_10pop_ex", ".png"), width = 13, height = 8, unit = "in", res = 225, antialias = "cleartype")

histplot5

dev.off()
```

### Animated Histograms

```{r animated_hist, fig.width = 20, fig.height = 10}
# Tip 6: Animation isn't so hard. In this case, it affords us the ability to look at a time lapse of sorts.

fp_output <- "~/GitHub/exercise-histogram-visualization/output/"


# Let's organize by region (roughly)
stateid$statename                                                               # getting the order of states, to add region identifiers.

# a manual list, I couldn't think of a better way for you to follow along! 
regions <-data.frame(c("South","West", "West","South","West","West","Northeast","Northeast","Northeast","South","South","West","West","Midwest","Midwest","Midwest","Midwest","South","South","Northeast","Northeast","Northeast","Midwest","Midwest","South","Midwest","West","Midwest","West","Northeast","Northeast","West","Northeast","South","Midwest","Midwest","South","West","Northeast","Northeast","South","Midwest","South","South","West","Northeast","South","West","Midwest","Midwest","West"))

# The default column name is annoying, let's change it.
colnames(regions) <- "region"

# gluing together region ids
stateid <- cbind(stateid, region = regions$region)                              # cbind() is used to combine columns of data frames. Here,
                                                                                # renaming the region column at the same time.

#New, compact data frame for our regional look.
regional <- left_join(state_wk, stateid, by = "statefips")                      # rebinding our identifiers like before, but now with an
                                                                                # added region identifier.

regional_wk <- dplyr::select(regional, 
                             date,                                              # select() returns all rows of the selected 
                             statename = statename.x,                           # the columns of a data frame.
                             region, 
                             gps_residential)  

#filter a new, larger date range.                                               # filter() and select() are powerful tools together!.
regional_test <- filter(regional_wk, between(isoweek(date), 10, 31))            # Notice I am filtering even more weeks now, from Mar 3 to
                                                                                # Jul 31 

# making the columns needed for the IQR geom_rect().
regional_test <- group_by(regional_test, region, isoweek(date)) %>%             #
  mutate(med = median(gps_residential)) %>%                                     # The last mutate() adds a column we haven't seen
  mutate(IQR1 = quantile(gps_residential, probs = 0.25)) %>%                    # check the grouping done. We find that
  mutate(IQR2 = quantile(gps_residential, probs = 0.75)) %>%                    # some regions have different observations due to
  mutate(n = length(gps_residential))                                           # having a different number of member states...                          

# check resulting data frame.                                                   # we need to use a density histogram. More on this later.
head(regional_test)                                                             # first, ugly column names!

colnames(regional_test) <- c("date", "state", "region", "time_in_home", "week", "med", "IQR25", "IQR75", "n")

# build the animation by layer.
anim <- ggplot(data = regional_test, aes(x=time_in_home, fill=region)) +        # note: changed column titles in previous step.
  
  scale_x_continuous(labels = scales::percent) + 
  scale_y_continuous(labels = scales::percent) +                                # scale both axis as continuous percent now.
  geom_rect(data = regional_test, inherit.aes = FALSE,                          # need to use the less ideal method of IQR highlighting
            aes(xmin = IQR25, xmax = IQR75, ymin=0, ymax=0.5, group = region),  
            fill = "#8496E4", alpha=0.00589) +                                  # trial and error fill and alpha values.
# note: now a density histogram
  geom_histogram(aes(y=(..density..)*0.015), binwidth = 0.015,                  # because there are a different number of records between,
                 alpha = 0.5) +                                                 # regions, we must add aes(y=..density..) to our histogram.
                                                                                # we must set out binwidth here, and multiply ..density.. by
                                                                                # our binwidth to rescale the y axis.
                                                                                # this will report percentage of total count, rather than
                                                                                # count.
  geom_vline(aes(xintercept = med, group = region), 
             color = "red", linetype = "dashed", size = 1, alpha = 0.8) +  
  
  facet_wrap(~ region, ncol=4) +                                                # changed number of columns on display.
  
  geom_rug(aes(color = region)) + 
 
  theme_minimal() +                                                                                                                                                
  theme(legend.position = "none") +                      
                                                                                # regarding ggtitle below!
  theme(strip.text = element_text(face="bold", size=12,lineheight=5.0)) +       # \n Week: {closest_state} will display week
                                                                                # because we will map our animation transition to week.
  ggtitle("Weekday Hours at Home By U.S. Region, 3/2/20 - 7/31/20 Weekly Snapshot \n Week: {closest_state}",               
        subtitle = "weekday avgs (by day) for all u.s. states, grouped by region, compared to jan '20 median levels" ) +          
        
  xlab("Daily Changes in Hours Home (Each Weekday vs. Jan Median)") +
  
  ylab("Percent of Total Days") +                                               # change y axis label, due to density histogram
  
  transition_states(week, transition_length = 1, state_length = 8)              # transition_states() turns the plot into an animation
                                                                                # the simplest way to define an animation is shown.
rendered.animation <- gganimate::animate(anim,                                             # Think of state_length() and transition_length() as a
                              duration = 60,                                    # ratio of the total duration set in animate().
                              fps = 24,                                         # Careful, large values here can cause the animation to take
                              height = 900,                                     # a long time to render!
                              width = 1600,
                              start_pause = 3,                                  # small pause of first week
                              end_pause = 3,                                    # small pause on second week
                              rewind = TRUE                                     # play the animation backwards, at the end.
                              )              

anim_save(filename = paste0(fp_output, "timeinhomehistograms.gif"), 
          animation = rendered.animation)
```
##Top Ten Things I Think are Worth Remembering

1) Filter()/Select()/Group_by()

2) Always Add Alpha

3) Medians and IQRs

4) The default, Pretty(Sturges) bin size function.  Also, binwidth = weirdbins!

5) Rugs and color grouping

6) Small Multiples with facets (theme_minimal())

7) Building a basic animation with gganimate

8) Different counts? Density histogram!

9) Adding text with geom_text(), and other major labels to ggplots. 

10) Histograms aren't bar charts!
